\bf Model Visualization Techniques for a Social Network Model

### 2017-04-11 paper.Rnw
bit of sec 4.2

\section{Methods for Network Visualization}

@@

Some layout algorithms were designed to mimic physical systems, drawing the graphs based on the "forces" connecting them. The network's edges act as springs pushing and pulling the nodes in 2D space. Others use properties of the network's adjaceny matrix in order to place the nodes. Some commonly used layout algorithms are  
\begin{itemize}
\item Kamada-Kawai: The ``graph theoretic distance'' from one node to another is represented in the 2D node placement. This network drawing method results in several nice properties such as ``symmetric drawings [and] a relatively small number of edge crossings'' \citep[p. 15]{kamadakawai}.
\item Fructerman-Reingold: This algorithm is a spring-embedder type where ``the vertices behave as [...] celestial bodies, exerting attractive and repulsive forces on one another,'' where the simple goal of drawing connected nodes near to each other \citep[p. 1131]{fruchterman-reingold}.
\item Multi-dimensional scaling:
\item Eigen
\end{itemize}

@@

# running Siena
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(drink[20:35,])
mydata <- sienaDataCreate(friendship, alcohol)
myeffnull <- getEffects(mydata)
myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
ansnull <- siena07(myalgorithm, data = mydata, effects = myeffnull,
                   returnChains = TRUE, returnDataFrame = TRUE,
                   returnDeps = TRUE, silent = TRUE, verbose = FALSE,
                   batch = TRUE)
ansnullchains <- get_chain_info(ansnull)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:9]

@@

In \citeauthor{modelvis}, they define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. What does this definition mean for network data? For dynamic social networks, there are a few different data ``spaces." First, we have the actors and their corresponding covariates. Second, we have the dyad variables that describe the relationships between the variables. Finally, we have the time over which the network is allowed to evolve. These three data pieces can be visualized together in various ways. The traditional network visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if the dyad between two nodes has value greater than zero, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side. 
\st{
\begin{itemize}
\item WHat is the model? the network model (SAOM model is just one of many)
\item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
\item Note: specifically dealing with dynamic networks observed at discrete time points. 
\end{itemize}
What are some things that could go in here? need more data examples; not just the friends data.
Harry Potter data??? I want to find my own data, too.}

>>>>

\section{Networks and their Visualizations}

\subsection{Introduction to Network Structures}

\subsection{Visualizing Network Data} \label{sec:netvizintro}

@@

Some layout algorithms were designed to mimic physical systems, drawing the graphs based on the "forces" connecting them. The network's edges act as springs pushing and pulling the nodes in 2D space. Others use properties of the network's adjaceny matrix in order to place the nodes. Some commonly used layout algorithms are  
\begin{itemize}
\item Kamada-Kawai: The ``graph theoretic distance'' from one node to another is represented in the 2D node placement. This network drawing method results in several nice properties such as ``symmetric drawings [and] a relatively small number of edge crossings'' \citep[p. 15]{kamadakawai}.
\item Fructerman-Reingold: This algorithm is a spring-embedder type where ``the vertices behave as [...] celestial bodies, exerting attractive and repulsive forces on one another,'' where the simple goal of drawing connected nodes near to each other \citep[p. 1131]{fruchterman-reingold}.
\item Multi-dimensional scaling: 
\item Eigen:
\end{itemize}

@@

# running Siena
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(drink[20:35,])
mydata <- sienaDataCreate(friendship, alcohol)
myeffnull <- getEffects(mydata)
myalgorithm <- sienaAlgorithmCreate(projname = 's16_3')
set.seed(823746)
load("data/ansnull.rda")
ansnullchains <- get_chain_info(ansnullpaper)
ansnullchains %>%
  filter(period == 1) %>%  #only look at chains from wave 1 to wave 2
  group_by(rep) %>%
  select(rep, from = ego, to = alter) %>%
  mutate(val = as.numeric(!from == to),
         from = paste0("V", parse_number(from)+1), # make the chains
         to = paste0("V", parse_number(to)+1)) -> ansnullchainsw1w2
colnames(fd2.w1) <- paste0("V", 1:16)
rownames(fd2.w1) <- paste0("V", 1:16)
wave1friends <- fortify(as.adjmat(fd2.w1))
ms1 <- listMicrosteps(dat = wave1friends,
                      microsteps = filter(ansnullchainsw1w2, rep == 1))
microsteps <- ms1[1:9]

@@

In \citeauthor{modelvis}, they define the \textit{data space} as ``the region over which we can reliably make inferences, usually a hypercube containing the data" \citep[p.~206]{modelvis}. What does this definition mean for network data? For dynamic social networks, there are a few different data ``spaces." First, we have the actors and their corresponding covariates. Second, we have the dyad variables that describe the relationships between the variables. Finally, we have the time over which the network is allowed to evolve. These three data pieces can be visualized together in various ways. The traditional node-link visualization uses one of many algorithms to layout the actors as points in 2D space, then draws segments connecting the points in 2D if the dyad between two nodes has value greater than zero, and draws nothing otherwise. The time aspect can be visualized by drawing each network observation in time and placing the observed timepoints side-by-side. These visual connections between the three data spaces are themselves part of the data space, because they can be viewed in different ways. Another way to connect the node and the edge data, as discussed in Section~\ref{sec:netvizintro}, is through the network's adjaceny matrix. The time dimension can also be incorporated with side-by-side adjacency matrix visualizations. 
\st{
\begin{itemize}
\item WHat is the model? the network model (SAOM model is just one of many)
\item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
\item Note: specifically dealing with dynamic networks observed at discrete time points. 
\end{itemize}
What are some things that could go in here? need more data examples; not just the friends data.
Harry Potter data??? I want to find my own data, too.}
